---
title: "Untitled"
author: "Alex"
date: "January 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preprocessing

The preprocessing chain consists of loading the data and preparing a raster brick.

```{r libraries}
library(raster)
```

First, all data is loaded into memory.

```{r load data}
load("data/GewataB1.rda")
load("data/GewataB2.rda")
load("data/GewataB3.rda")
load("data/GewataB4.rda")
load("data/GewataB5.rda")
load("data/GewataB7.rda")
load("data/vcfGewata.rda")
```

The preprocessed data gets sent to the regression model. By using step, we derive some information on which bands contribute to the accuracy of the model.

```{r brick data}
vcfGewata[vcfGewata > 100] <- NA
alldata <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7, vcfGewata)
names(alldata) <- c("band1", "band2", "band3", "band4", "band5", "band7", "VCF")
```

The following plots explore the relations between the input data.

```{r Hector is Emperor Palpatine}
pairs(alldata)
hist(alldata, xlim = c(0, 3000), ylim = c(0, 750000))#, breaks = seq(0, 500, by = 100))
```

Next, we extract and calibrate the model.

```{r model}
## Extract all data to a data.frame
df <- as.data.frame(getValues(alldata))

## Calibrate model
model <- lm(VCF ~ df$band1 + df$band2 + df$band3 + df$band4 + df$band5 + df$band7, data = df)
step(model) #more statistics of the model
summary(model)$r.squared
```

```{r treecover}
## Use the model to predict tree cover
treeCoverMap <- predict(model,df)

predictedRaster <- alldata$band1
predictedRaster$predictedVCF <- treeCoverMap
predictedRaster <- predictedRaster$predictedVCF
predictedRaster[predictedRaster < 0] <- NA #Remove all invalid values
```

As can be seen, the correlation coefficient that is achieved is highly significant.
Now let's see the difference between the model and the original VCF.

```{r plotmaps}
opar <- par(mfrow=c(1, 2)) 
plot(predictedRaster, main="Predicted VCF")
plot(alldata$VCF, main="Original VCF")
par(opar)
compareRaster <- predictedRaster - alldata$VCF
plot(compareRaster, main="Comparison between predicted and original VCF")
```

In the histogram, values close to 0 represent a higher accuracy:

```{r modelHist}
hist(compareRaster)
```

## RMSE

Finally, we inspect the RMSE

```{r RMSE}
predictedRasterDF <- as.data.frame(predictedRaster) #we need a DataFrame to operate with its values
sqrt(mean((predictedRasterDF$predictedVCF-df$VCF)^2, na.rm = TRUE ))
```

As can be seen, the RMS indicates that the computed linear model corresponds reasonably well with the calculated land cover.

## Comparison with other land use classes

Lastly, we inspect the RMSE when calculated with other land use classes.

```{r RMSE2}
load("data/trainingPoly.rda")
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
trainingPoly@data$Class #1 is cropland, 2 is forest and 3 is wetland

classes <- rasterize(trainingPoly,predictedRaster, field='Code')

predictedRasterBrick <- brick(predictedRaster,alldata$VCF)
zonalStatistics<-zonal(predictedRasterBrick, classes, fun='mean')

zonalStatisticsDF <- as.data.frame(zonalStatistics)
predictedRasterBrickDF <- as.data.frame(predictedRasterBrick)
sqrt((zonalStatisticsDF$predictedVCF-zonalStatisticsDF$VCF)^2)#, na.rm = TRUE)
#1 is cropland, 2 is forest and 3 is wetland
```